---
output:
  bookdown::pdf_document2:
    fig_caption: yes
    latex_engine: xelatex
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
classoption: landscape
---
## England {#england}

```{r, include=FALSE}
#---------------------------------------
# knitr -- create pretty tables for a presentation
#---------------------------------------
if(!require(knitr)){install.packages("knitr")}
library(knitr)
#---------------------------------------
# kableExtra -- create pretty tables for a presentation
#---------------------------------------
if(!require(kableExtra)){install.packages("kableExtra")}
library(kableExtra)
#---------------------------------------
# tidyr -- replace_na
#---------------------------------------
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
#---------------------------------------
# dplyr -- lag
#---------------------------------------
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
#---------------------------------------
# ggplot2 -- plots
#---------------------------------------
if(!require(ggplot2)){install.packages("ggplot2")}
library(ggplot2)
#---------------------------------------
# ggExtra -- ggMarginal
#---------------------------------------
if(!require(ggExtra)){install.packages("ggExtra")}
library(ggExtra)
#---------------------------------------
# visdat -- vis_miss
#---------------------------------------
if(!require(visdat)){install.packages("visdat")}
library(visdat)
#---------------------------------------
# naniar -- gg_miss_upset
#---------------------------------------
if(!require(naniar)){install.packages("naniar")}
library(naniar)
#---------------------------------------
# reshape2 -- melt
#---------------------------------------
if(!require(reshape2)){install.packages("reshape2")}
library(reshape2)
```

```{r, include=FALSE}
source("my_kable.r")
```

Look at the dataset's columns.
```{r, echo=FALSE}
england <- read.csv("/Users/travel_mechtal/Documents/UWE/Portfolio/nation_2022-03-10.csv", header = TRUE, sep = ",")

my_kable(colnames(england))
```
We are not going to look at the columns areaCode, areaName, areaType, because these columns have one unique value ($`r unique(england["areaCode"])`$ for areaCode, $`r unique(england["areaName"])`$ for areaName, $`r unique(england["areaType"])`$ for areaType), they are used for filtering on the website.

Rename columns and exclude unnecessary columns.

```{r, echo=FALSE}
england$Date <- as.Date(england$date, format = "%Y-%m-%d")

england <- england[order(england$Date),]

england = rename(england, 
       First = newPeopleVaccinatedFirstDoseByVaccinationDate
       ,Second = newPeopleVaccinatedSecondDoseByVaccinationDate
       ,Third = newPeopleVaccinatedThirdInjectionByVaccinationDate
         )

england <- england[c("Date", "First", "Second", "Third")]

kable(head(england))
```

I want to be sure that the file has statistics for every day.

- Add the column with difference between the date and the previous date within the table.
```{r, include=FALSE}
# replce na for the first date by the previous date for the first date
england$Diffdate <- england$Date - replace_na(lag(england$Date), min(england$Date) - 1)
```
The difference between dates is $`r unique(england["Diffdate"])`$. So, we have full statistics by date.

Do we have missing values?
```{r, echo=FALSE}
# count of missing values per column
kable(colSums(is.na(england)), col.names = NULL)
```

```{r, echo=FALSE, fig.height = 5, fig.width = 10, fig.align = "center"}
vis_miss(england)
```
Yes, we have missing values in the column "Third". In this case, missing values and zeroes are equivalent.

Replace by zero.

```{r, include=FALSE}
england$Third <- england$Third %>% replace_na(0)
```

```{r, include=FALSE}
myFirst <- as.Date("2021-08-08", format="%Y-%m-%d")
mySecond <- as.Date("2021-10-03", format="%Y-%m-%d")
myThird <- as.Date("2022-01-08", format="%Y-%m-%d")
```

```{r, include=FALSE}
england_long <- melt(england, id.vars = c("Date")
                       , measure.vars = c("First", "Second", "Third")
                       , variable.name = "Dose"
                       , value.name = "Count"
                    )
```

```{r, include=FALSE}
myData <- filter(england_long, Dose=="First" & Date==myFirst |
                     Dose=="Second" & Date==mySecond |
                     Dose=="Third" & Date==myThird )
```

```{r, include=FALSE}
maxData <- england_long %>%
  group_by(Dose) %>%
  top_n(1, Count)
```

I have a question.

```{r ref0, ref.label='q1', render=pander::pander, echo=FALSE}
```

```{r, echo=FALSE, fig.height = 8, fig.width = 12, fig.align = "center"}
ggplot(england_long, aes(x=Date, y=Count)) +
  geom_line() +
  geom_point(data=maxData) +
  geom_label(data=maxData, aes(label = Count, y = Count + 80000, colour="max")) +
  geom_point(data=myData, colour="red") +
  geom_label(data=myData, aes(label = Count, y = Count + 160000, colour="my")) +
  scale_x_date(date_breaks = "months" , date_labels = "%b-%y") +
  scale_y_continuous(labels = scales::comma) +
  facet_wrap(~ Dose, ncol=1) +
  theme(axis.text.x = element_text(angle=90, hjust = 1)) +
  scale_color_manual(name = "Points", values = c("max counts" = "black", "my dates" = "red")) +
  labs(x = "Date" 
    , y = "Count of doses"
    , title = "Vaccination in England"
  )
```