---
site: bookdown::bookdown_site
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: yes
    latex_engine: xelatex
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
classoption: landscape
---
# Exploring datasets

## South West {#DatasetSouthWest}

```{r DatasetSouthWestPaths, include=FALSE}
path = "/Users/travel_mechtal/Documents/UWE/Portfolio/"
```

```{r DatasetSouthWestLibraries, include=FALSE}
#---------------------------------------
# Knitting to pdf
#---------------------------------------
if(!require(knitr)){install.packages("knitr")}
library(knitr)
#---------------------------------------
if(!require(kableExtra)){install.packages("kableExtra")}
library(kableExtra)
#---------------------------------------
if(!require(tinytex)){install.packages("tinytex")}
library(tinytex)
#---------------------------------------
if(!require(pander)){install.packages("pander")}
library(pander) #beautiful string from chunk
#---------------------------------------
# Describing data
#---------------------------------------
if(!require(psych)){install.packages("psych")}
library(psych)
#---------------------------------------
# Plotting
#---------------------------------------
if(!require(ggplot2)){install.packages("ggplot2")}
library(ggplot2)
#---------------------------------------
# Working with objects
#---------------------------------------
if(!require(tidyverse)){install.packages("tidyverse")}
library(tidyverse) #work with objects using %>%, for example
#---------------------------------------
if(!require(reshape2)){install.packages("reshape2")}
library(reshape2) #use melt, for example
#---------------------------------------
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate) #work with dates
```

```{r DatasetSouthWestRFunctions, include=FALSE}
#---------------------------------------
# My function for a list view
#---------------------------------------
source("my_kable.R")
```

Look at the dataset for South West.

```{r DatasetSouthWest, echo=FALSE}
# read the dataset
region_set <- read.csv(paste0(path,"region_2022-01-27.csv"),
                         header = TRUE,
                         sep = ","
                         )

# change date's format
region_set$date <- as.Date(region_set$date,format = '%Y-%m-%d')

my_kable(colnames(region_set))
```


We have additional columns. Let's look at them.

- For **areaCode** unique value is `r unique(region_set[,"areaCode"])`,

- for **areaName** unique value is `r unique(region_set[,"areaName"])`,

- for **areaType** unique value is `r unique(region_set[,"areaType"])`.

So, we do not need to look at them in the future because these columns are used for filtering that we have already done on the website.

Let's prepare data for the plotting.

- Rename columns and columns

```{r DatasetSouthWestNewNames, include=FALSE}
# rename columns
region_set = rename(region_set, 
       First = newPeopleVaccinatedFirstDoseByVaccinationDate
       ,Second = newPeopleVaccinatedSecondDoseByVaccinationDate
       ,Third = newPeopleVaccinatedThirdInjectionByVaccinationDate
         )
```

- Create a long table

```{r DatasetSouthWestLong, include=FALSE}
# add the long verstion
region_set_long <- melt(region_set
                       , id.vars = c("date")
                       , measure.vars = c("First"
                                          , "Second"
                                          , "Third"
                                       )
                       , variable.name = "Dose"
                       , value.name = "Count"
                    )
```

```{r DatasetSouthWestLongHead, echo=FALSE}
kable(head(region_set_long))
```

Let's plot data.

```{r DatasetSouthWestVariables, include=FALSE}
# create variables for my dates
myFirst <- as.Date("2021-08-08", format="%Y-%m-%d")
mySecond <- as.Date("2021-10-03", format="%Y-%m-%d")
myThird <- as.Date("2022-01-08", format="%Y-%m-%d")
```

```{r DatasetSouthWestLongFilter, include=FALSE}
region_set_long_period <- filter(region_set_long, date >= as.Date("2021-07-01", format = "%Y-%m-%d"))
```

As I say in the part \@ref(Questions), I have a question:
```{r ref0, ref.label='q0', render=pander::pander, echo=FALSE}
```

Let's answer.
The red lines are my vaccination dates.
\newpage
```{r DatasetSouthWestAllDosesPlot, echo=FALSE}
ggplot(data = region_set_long_period
       , aes(x=date
             , y=Count
             , colour = Dose
             )
       ) + 
geom_point() +
theme(axis.text.x = element_text(angle=90, hjust = 1)) +
geom_vline(xintercept = myFirst, color = "red") +
geom_vline(xintercept = mySecond, color = "red") +
geom_vline(xintercept = myThird, color = "red") 
```
\newpage
The result is not beautiful because of the active growth of the third jabs count at the end of 2021. 

Let's plot them separately.

```{r DatasetSouthWestFirst, include=FALSE}
PlotDose <- "First"
```

```{r DatasetSouthWestPlot, echo=FALSE}
ggplot(data = filter(region_set_long_period, Dose == PlotDose)
       , aes(x=date
             , y=Count
             )
       ) + 
geom_point(colour = "blue") +
theme(axis.text.x = element_text(angle=90, hjust = 1)) +
geom_vline(xintercept = get(paste0("my", PlotDose)), color = "red") +
  labs(
    x = "Date"
    , y = paste("Count of", PlotDose, "doses")
    , title = "Vaccination in South West"
    , subtitle = paste("The", PlotDose, "dose")
    , caption = "More information https://coronavirus.data.gov.uk/details/about-data"
  )
```

It is so interesting why the graph is wavy. `r filter(region_set, date==myFirst)["First"]` people got their first jabs with me.

```{r DatasetSouthWestSecond, echo=FALSE}
PlotDose <- "Second"
<<DatasetSouthWestPlot>>
```

`r filter(region_set, date==mySecond)["Second"]` people got their second jabs with me.

```{r DatasetSouthWestThird, echo=FALSE}
PlotDose <- "Third"
<<DatasetSouthWestPlot>>
```

`r filter(region_set, date==myThird)["Third"]` people got their third jabs with me.
We can see when the active phase of vaccination by the third dose started.

\newpage
### Zeroes

```{r DatasetSouthWestZeroSet, echo=FALSE, warning=FALSE}
region_set_long <- region_set_long %>% mutate(zero =
           case_when(Count == 0 ~ "0", 
                     Count > 0 ~ "> 0")        
           )          
```

```{r DatasetSouthWestZeroPlot, echo=FALSE, fig.align = "center", fig.height = 3, fig.width = 12, message = FALSE}
ggplot(region_set_long %>%
        group_by(Dose, zero) %>%
        summarise(zero_count=length(Count))
, aes(x="", y=zero_count, fill=zero)) +
  geom_bar(stat="identity", width=10) +
  coord_polar("y", start=0) +
  facet_grid(cols = vars(Dose)) +
  theme_void() # remove background, grid, numeric labels
```

The column "Third" has more zero values than "First" and "Second; but, I think, it won't influence models' accuracy. 
Also, we can see missing values for the column "Third"; in our case, missing values mean that nobody got the third jab. I suggest replacing them with zeroes.

Replace missing values.

```{r DatasetSouthWestLongNA, include=FALSE}
region_set_long[is.na(region_set_long)] <- 0
```

\newpage
### Data description

```{r DatasetSouthWestDesc, include=FALSE}
dataset_desc <- describe(region_set[,c("First", "Second", "Third")], IQR=TRUE, quant=c(.25,.75,.90))
```

_Median, percentiles and mean_

```{r DatasetSouthWestDescPart1, echo=FALSE, eval=TRUE, echo=FALSE}
kable(dataset_desc[, c("mean", "median", "Q0.25", "Q0.75", "Q0.9")])
```

What can I say?

- Mean and median have a visible difference. So, there are large extreme values.

- For the Third dose, half of the values are below $`r dataset_desc["Third", "median"]`$. That is not surprised. In the beginning, people needed to get two jabs.

- If we look at "Q0.25", "Q0.75", "Q0.90", we find out that the Third dose's wave caught up with other doses' waves quickly. We already saw this fact on the plots.

```{r DatasetSouthWestMeanMedianDataset, include=FALSE}
a_mean <- region_set_long %>% 
  group_by(Dose) %>% 
  summarize(mean_val = mean(Count),
            median_val = median(Count))
```

```{r DatasetSouthWestHistogram, echo=FALSE, fig.align = "center"}
region_set_long %>%
ggplot(aes(x=Count, fill=Dose)) +
  geom_histogram(color="#e9ecef", bins = 15, position = 'identity') +
  facet_wrap(~ Dose, ncol=1) +
  geom_vline(data= a_mean, aes(xintercept=mean_val, color="mean"), linetype="dashed", size=1) +
  geom_vline(data= a_mean, aes(xintercept=median_val, color="median"), linetype="dashed", size=1) +
  scale_color_manual(name = "Statistics", values = c("mean" = "blue", "median" = "red")) 
```
_Standard deviation (sd), IQR and range_

```{r DatasetSouthWesxtDescPart2, echo=FALSE, eval=TRUE, echo=FALSE}
kable(dataset_desc[,c("sd", "range", "IQR")])
```
IQR and standard deviation for each dose are big, consequently, the data spread out.
Also, we can see the difference between largest and smallest values in the column "range".