---
output:
  bookdown::pdf_document2:
    fig_caption: yes
    latex_engine: xelatex
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
classoption: landscape
---
# Choosing datasets

```{r DatasetsPaths, include=FALSE}
path = "/Users/travel_mechtal/Documents/UWE/Portfolio/"
```

```{r DatasetsLibraries, include=FALSE}
#---------------------------------------
# Knitting to pdf
#---------------------------------------
if(!require(tinytex)){install.packages("tinytex")}
library(tinytex)
#---------------------------------------
# Creating pretty tables
#---------------------------------------
if(!require(knitr)){install.packages("knitr")}
library(knitr)
#---------------------------------------
if(!require(kableExtra)){install.packages("kableExtra")}
library(kableExtra)
#---------------------------------------
# Working with objects
#---------------------------------------
# json
if(!require(jsonlite)){install.packages("jsonlite")}
library(jsonlite)
#---------------------------------------
# hash
if(!require(hash)){install.packages("hash")}
library(hash)
#---------------------------------------
# %>%
if(!require(tidyverse)){install.packages("tidyverse")}
library(tidyverse)
```

```{r DatasetsRFunctions, include=FALSE}
#---------------------------------------
# My function for a list view
#---------------------------------------
source("my_kable.R")
#---------------------------------------
# My function to show which metrics belong to the dataset
#---------------------------------------
source("show_links.R")
```

Look at the vaccination metrics from the website ["Coronavirus (COVID-19) in the UK"](https://coronavirus.data.gov.uk/metrics/category/Vaccinations).

```{r DatasetsVaccinationMetricsAll, echo=FALSE}
# transform to dataframe json file
VaccinationMetrics <- subset(as.data.frame(read_json(paste0(path, "metrics_Vaccinations.json"), simplifyVector = TRUE)), category == "Vaccinations")

my_kable(VaccinationMetrics$metric_name)
```

I am going to use metrics 

1. which names start with “New”, 
2. by vaccination date, 
3. and are not deprecated.

Which [datasets](https://coronavirus.data.gov.uk/metrics/area_type) have these metrics?

```{r DatasetsVaccinationMetrics, echo=FALSE}
# create a set of necessary metrics
VaccinationMetricsMy <- subset(VaccinationMetrics,
  # conditions for subset
  startsWith(metric_name, "New")
  & endsWith(metric_name, "by vaccination date")
  & is.na(deprecated)
  )[,c("metric", "metric_name")] #columns

# create sets of metrics for the first, the second, and the third doses
first = VaccinationMetricsMy$metric[grepl("1st", VaccinationMetricsMy$metric_name, fixed = TRUE) | grepl("first", VaccinationMetricsMy$metric_name, fixed = TRUE)]

second = VaccinationMetricsMy$metric[grepl("2nd", VaccinationMetricsMy$metric_name, fixed = TRUE) | grepl("second", VaccinationMetricsMy$metric_name, fixed = TRUE) 
                     ]
third = VaccinationMetricsMy$metric[grepl("third", VaccinationMetricsMy$metric_name, fixed = TRUE) | grepl("booster", VaccinationMetricsMy$metric_name, fixed = TRUE) 
                     ]
```

```{r DatasetsAreaTypes, include=FALSE}
# create dictionary: short area type's name = area type's name
AreaTypesDict <- hash("ltla"="Lower Tier Local Authority (LTLA)"
                      ,"msoa"="Middle-layer Super Output Area (MSOA)"
                      ,"nation"="Nation"
                      ,"nhsRegion"="NHS Region"
                      ,"nhsTrust"="NHS Trust"
                      ,"overview"="Overview"
                      ,"region"="Region"
                      ,"utla"="Upper Tier Local Authority (UTLA)"
                      )

for(AreaType in keys(AreaTypesDict)){
  # AreaType metrics' file path
  full_path <- paste0(path, AreaType, "_new_vaccination")
  # create vector with metrics and name it as AreaType's value
  assign(AreaType, scan(full_path, character(), quote = ""))
}
```

```{r DatasetsMatrixM, include=FALSE}
# create matrix to show which metrics belong to the datasets
M = matrix(0, length(VaccinationMetrics$metric), length(keys(AreaTypesDict)))
rownames(M) = VaccinationMetrics$metric
colnames(M) = keys(AreaTypesDict)
# show links
for(AreaType in keys(AreaTypesDict)){
M = show_links(M, AreaType)
}
```

```{r DatasetsMatrixMView, echo=FALSE}
# view the table with highlighted necessary metrics
M %>%
  kable(caption = "Vaccination metrics with highlight") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  column_spec(1, width = "30em") %>%
  row_spec(which(VaccinationMetrics$metric %in% first), color = "#104e8b", background = "#d3d3d3") %>%
  row_spec(which(VaccinationMetrics$metric %in% second), color = "red", background = "#d3d3d3") %>%
  row_spec(which(VaccinationMetrics$metric %in% third), color = "black", background = "#d3d3d3")
```

So, I need to look at the datasets:

```{r DatasetsFirst, echo=FALSE}
# choose the datasets which contain necessary metrics
FirstDataset <- c()

for(AreaType in keys(AreaTypesDict)){
  for(row in first) {
    if(M[row,AreaType] ==1)
      {FirstDataset=append(FirstDataset,AreaTypesDict[[AreaType]])}
  }
}

my_kable(unique(FirstDataset))
```

The dataset "Overview" presents the UK.
Which metrics do we have for this dataset?
```{r DatasetsUK, echo=FALSE}
my_kable(overview)
```

So, we do not have data by vaccination date for the UK. I am going to use metrics from the Table \@ref(tab:DatasetsMatrixMView) for

1. Bristol (Lower Tier Local Authority (LTLA)),
2. South West (Region)
3. and England (Nation).

That is enough for answering.
