---
output:
  bookdown::pdf_document2:
    fig_caption: yes
    latex_engine: xelatex
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
classoption: landscape
---
## England {#ModelEngland}

```{r ModelEnglandPaths, include=FALSE}
path = "/Users/travel_mechtal/Documents/UWE/Portfolio/"
```

```{r ModelEnglandLibraries, include=FALSE}
#---------------------------------------
# Knitting to pdf
#---------------------------------------
if(!require(knitr)){install.packages("knitr")}
library(knitr)
#---------------------------------------
if(!require(kableExtra)){install.packages("kableExtra")}
library(kableExtra)
#---------------------------------------
# Working with objects
#---------------------------------------
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr) #rename
#---------------------------------------
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr) #replace_na
#---------------------------------------
if(!require(glue)){install.packages("glue")}
library(glue) #glue_collapse
#---------------------------------------
# Forcasting
#---------------------------------------
if(!require(forecast)){install.packages("forecast")}
library(forecast)
#---------------------------------------
# Plotting
#---------------------------------------
if(!require(ggplot2)){install.packages("ggplot2")}
library(ggplot2)
```

Look at the dataset.

```{r ModelEnglandDataset, echo=FALSE}
# read the data
england <- read.csv(paste0(path,"nation_2022-03-10.csv"), header = TRUE, sep = ",")
# add new column Date
england$Date <- as.Date(england$date, format = "%Y-%m-%d")
# order the dataset
england <- england[order(england$Date),]
# rename columns
england = rename(england, 
       First = newPeopleVaccinatedFirstDoseByVaccinationDate
       ,Second = newPeopleVaccinatedSecondDoseByVaccinationDate
       ,Third = newPeopleVaccinatedThirdInjectionByVaccinationDate
         )
# choose necessary columns
england <- england[c("Date", "First", "Second", "Third")]
# replace NA-values for the column "Third"
england$Third <- england$Third %>% replace_na(0)
# add the new column All that equals a sum of doses counts
england$All <- england$First + england$Second + england$Third

kable(head(england))
```

Look at the plot for the sum of doses counts.

```{r ModelEnglandAll, include=FALSE}
FitColumn <- "All"
```

```{r ModelEnglandPlot, echo=FALSE}
ggplot(england, aes_string(x="Date", y=FitColumn)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma)
```

What can we say?

There is a not stationary time series, as the series wanders up and down for long periods.

I am going to use an ARIMA model.

```{r ModelEnglandTimeSeries, include=FALSE}
englandFit <- ts(data=england[FitColumn], frequency = 1)
fit <- auto.arima(englandFit)
pdq <- arimaorder(fit)
```

ARIMA(`r glue_collapse(pdq, sep=", ")`)

* `r pdq[1]` – is the order of Auto-regressive or linear model
* `r pdq[2]` – difference value to make the time series stationary from non-stationary.
* `r pdq[3]` – is the order of Moving Average/ number of lagged values

If the data is stationary, then d=0. So, I was right earlier.

```{r ModelEnglandTimeSeriesAcf, echo=FALSE}
acf(englandFit)
```

As we know, the autocorrelation function (ACF) assesses the correlation between observations in a time series for a set of lags. In an ACF plot, each bar represents the size and direction of the correlation. Bars that extend across the blue line are statistically significant.

So, 

* this ACF plot indicates that these time series data are not random.
* the autocorrelations decline slowly.
* When a time series has both a trend and seasonality, the ACF plot displays a mixture of both effects. Notice how you can see the wavy correlations for the seasonal pattern and the slowly diminishing lags of a trend.

Look at the residuals that tell us a story about the model performance.

```{r ModelEnglandArimaCheckResiduals, echo=FALSE, message=FALSE}
checkresiduals(fit)
```

The first plot shows us that the difference between the observations and the corresponding fitted values is not fine.

```{r ModelEnglandArimaForecast, echo=FALSE}
autoplot(forecast(fit))
```

I am not impressed because there are negative values. But I am sure that this may be corrected by using not auto.arima. What do I need to think about? Is how to explain to the model that the second dose is two months after the First, and the Third one is three or six months after the Second. Maybe I can find the explanatory variable for that.
