---
title: "Coronavirus (COVID-19) vaccination in the UK"
author: "Elena Basargina"
date: "14/02/2022"
output:
  beamer_presentation:
    theme: "Antibes"
    colortheme: "seahorse"
fontsize: 6pt
---
```{r PresenationLibraries, include=FALSE}
#---------------------------------------
# ggplot2 -- plot
#---------------------------------------
if(!require(ggplot2)){install.packages("ggplot2")}
library(ggplot2)
#---------------------------------------
# dplyr -- rename
#---------------------------------------
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
#---------------------------------------
# knitr -- create pretty tables for a presentation
#---------------------------------------
if(!require(knitr)){install.packages("knitr")}
library(knitr)
#---------------------------------------
# reshape2 -- use melt
#---------------------------------------
if(!require(reshape2)){install.packages("reshape2")}
library(reshape2)
#---------------------------------------
# lubridate -- work with dates
#---------------------------------------
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
#---------------------------------------
# psych -- describing data
#---------------------------------------
if(!require(psych)){install.packages("psych")}
library(psych)
```

```{r python, include=FALSE}
# Setting for using python
library(tidyverse)
library(reticulate)
use_python("/Users/travel_mechtal/opt/anaconda3/bin/python")
use_condaenv("base", required = TRUE)

# if I need a new module
#py_install("ModuleName")
# I already installed: pandas, networkx, typing_extensions, joblib, seaborn, scikit-learn
```

```{python python_libraries, include=FALSE}
# work with dataframes
import pandas as pd
# work with dates
import datetime as dt
# split the dataset
from sklearn.model_selection import train_test_split
# evaluate the model
from sklearn.metrics import mean_absolute_error
# graph data
pd.plotting.register_matplotlib_converters()
import matplotlib.pyplot as plt
import seaborn as sns
from sklearn import tree
from sklearn.tree import export_text
# build a model
from sklearn.tree import DecisionTreeRegressor
from sklearn.ensemble import RandomForestRegressor
```

# Introduction
## What do I know about Bristol?

- This city is a part of the UK, England, and South West. 

- There are two universities.

# Data

I use the dataset from the 
[website](https://coronavirus.data.gov.uk/details/vaccinations) about Coronavirus (COVID-19) in the UK.

So, I am interested in data about the UK, England, South West, and Bristol. 

```{r, include=FALSE}
path = "/Users/travel_mechtal/Documents/UWE/Portfolio/"
```

**Bristol**
```{r, echo=FALSE}
bristol <- read.csv(paste(path,"ltla_2022-02-10.csv",sep=""),
                         header = TRUE,
                         sep = ","
                         )

# change date's format
bristol$date <- as.Date(bristol$date,format = '%Y-%m-%d')

bristol = rename(bristol, 
       First = newPeopleVaccinatedFirstDoseByVaccinationDate
       ,Second = newPeopleVaccinatedSecondDoseByVaccinationDate
       ,Third = newPeopleVaccinatedThirdInjectionByVaccinationDate
         )

bristol <- bristol[c("age", "date", "First", "Second", "Third")]

bristol[is.na(bristol)] <- 0

kable(head(bristol[1,]))
```
```{r, include=FALSE}
# add the long verstion
bristol_long <- melt(bristol
                       , id.vars = c("date",  "age")
                       , measure.vars = c("First"
                                          , "Second"
                                          , "Third"
                                       )
                       , variable.name = "Dose"
                       , value.name = "Count"
                    )
```

**South West**
```{r, echo=FALSE}
southwest <- read.csv(paste(path,"region_2022-01-27.csv",sep=""),
                         header = TRUE,
                         sep = ","
                         )

# change date's format
southwest$date <- as.Date(southwest$date,format = '%Y-%m-%d')

southwest = rename(southwest, 
       First = newPeopleVaccinatedFirstDoseByVaccinationDate
       ,Second = newPeopleVaccinatedSecondDoseByVaccinationDate
       ,Third = newPeopleVaccinatedThirdInjectionByVaccinationDate
         )

southwest <- southwest[c("date", "First", "Second", "Third")]

southwest[is.na(southwest)] <- 0

kable(head(southwest[1,]))
```
```{r, include=FALSE}
# add the long verstion
southwest_long <- melt(southwest
                       , id.vars = c("date")
                       , measure.vars = c("First"
                                          , "Second"
                                          , "Third"
                                       )
                       , variable.name = "Dose"
                       , value.name = "Count"
                    )
```

```{python read_data, include=FALSE}
# read file
path = "/Users/travel_mechtal/Documents/UWE/Portfolio/"
data=pd.read_csv(path + "region_2022-01-27.csv", 
index_col="date", parse_dates=True)

# drop unnecessary columns: areaCode, areaName, areaType.
dataset = data.drop(["areaCode", "areaName", "areaType"], axis=1)

# rename columns
# newPeopleVaccinatedFirstDoseByVaccinationDate -> First, 
# newPeopleVaccinatedSecondDoseByVaccinationDate -> Second, 
# newPeopleVaccinatedThirdInjectionByVaccinationDate -> Third
dataset = dataset.rename(columns={"newPeopleVaccinatedFirstDoseByVaccinationDate":"First", 
                                  "newPeopleVaccinatedSecondDoseByVaccinationDate":"Second", 
                                  "newPeopleVaccinatedThirdInjectionByVaccinationDate":"Third"})
# Replace Na values                                  
dataset = dataset.fillna(0) 
```


# What do I know about the data?

```{r, include=FALSE}
bristol_desc <- describe(bristol[,c("First", "Second", "Third")], IQR=TRUE, quant=c(.25,.75,.90))
```

```{r, include=FALSE}
southwest_desc <- describe(southwest[,c("First", "Second", "Third")], IQR=TRUE, quant=c(.25,.75,.90))
```

- There may be incorrect data

- There are missing values

- The counts of doses per day spread out. 

- There are sets of small and large counts.

# Questions 1

```{r, include=FALSE}
myFirst <- as.Date("2021-08-08", format="%Y-%m-%d")
mySecond <- as.Date("2021-10-03", format="%Y-%m-%d")
myThird <- as.Date("2022-01-08", format="%Y-%m-%d")
myAge <- "25_29"
```

I was vaccinated by 

- the first dose on $`r myFirst`$,

- the second dose on $`r mySecond`$, 

- the booster dose on $`r myThird`$.

How many people got their jabs with me?

---

## Bristol

```{r, echo=FALSE}
kable(filter(bristol_long, ((date == myFirst & Dose == "First") | 
                            (date == mySecond & Dose == "Second") |
                            (date == myThird & Dose == "Third")
                            ) 
             & age == myAge))
```

```{r, include=FALSE}
dose <- "First"
```

---

```{r BristolAgePlot,  fig.height = 6.9, fig.width = 11.5, fig.align = "center", echo=FALSE}
ggplot(data = bristol, 
       aes_string(x="date", 
           y=dose,
           colour="age")
       ) +
  geom_line() +
  facet_grid(age ~ .) +
  geom_vline(xintercept = get(paste("my", dose, sep='')), color = "red") +
  theme(strip.text.y = element_text(angle = 0)) +
  scale_x_date(date_breaks = "months" , date_labels = "%b-%y") +
  scale_y_continuous(n.breaks = 3)
```

---

```{r, fig.height = 6.9, fig.width = 11.5, fig.align = "center", echo=FALSE}
dose <- "Second"
<<BristolAgePlot>>
```

---

```{r, fig.height = 6.9, fig.width = 11.5, fig.align = "center", echo=FALSE}
dose <- "Third"
<<BristolAgePlot>>
```

---

## South West

```{r, echo=FALSE}
kable(filter(southwest_long, ((date == myFirst & Dose == "First") | 
                            (date == mySecond & Dose == "Second") |
                            (date == myThird & Dose == "Third")
                            ) 
             ))
```

```{r, include=FALSE}
# add columns
MonthYear <- c(c(paste(1:12,2020,sep=".")), c(paste(1:12,2021,sep=".")), c(paste(1:12,2022,sep=".")))
southwest$MonthYear <- factor(paste(month(southwest$date), year(southwest$date), sep=".")
                               , levels = MonthYear)
```

```{r, include=FALSE}
# add the long verstion
region_set_long <- melt(southwest
                       , id.vars = c("date", "MonthYear")
                       , measure.vars = c("First"
                                          , "Second"
                                          , "Third"
                                       )
                       , variable.name = "Dose"
                       , value.name = "Count"
                    )

region_set_long_period <- region_set_long #filter(region_set_long, date >= as.Date("2021-07-01", format = "%Y-%m-%d"))
```


```{r, include=FALSE}
dose <- "First"
```

---

```{r SouthWestPlot,fig.height = 6.9, fig.width = 11.5, fig.align = "center", echo=FALSE}
ggplot(data = filter(region_set_long_period, Dose == dose)
       , aes(x=date
             , y=Count
             )
       ) + 
geom_point(colour = "blue") +
theme(axis.text.x = element_text(angle=90, hjust = 1)) +
geom_vline(xintercept = get(paste("my", dose, sep='')), color = "red") +
    scale_x_date(date_breaks = "months" , date_labels = "%b-%y") +
  labs(
    x = "Date"
    , y = paste("Count of", dose, "doses")
    , title = "Vaccination in South West"
    , subtitle = paste("The", dose, "dose")
    , caption = "More information https://coronavirus.data.gov.uk/details/about-data"
  )
```

---

```{r, fig.height = 6.9, fig.width = 11.5, fig.align = "center", echo=FALSE}
dose <- "Second"
<<SouthWestPlot>>
```

---

```{r, fig.height = 6.9, fig.width = 11.5, fig.align = "center", echo=FALSE}
dose <- "Third"
<<SouthWestPlot>>
```

---

# Conclusion

I got my jabs in the not active phases of the vaccination, but I got the third one with a lot of people.

Also, I can say that there are waves. So, the count of jabs depends on dates. Keep this fact in mind.

# Question 2

When do people prefer to get a jab: weekdays or weekends/Saturdays or Sundays? 

---
```{python DateFeatures, include=FALSE}
dataset['Year'] = dataset.index.year
dataset['Month'] = dataset.index.month
dataset['Day'] = dataset.index.day
dataset['DayOfYear'] = dataset.index.dayofyear
dataset['Weekday'] = dataset.index.weekday
dataset['Quarter'] = dataset.index.quarter
dataset['IsMonthStart'] = dataset.index.is_month_start
dataset['IsMonthEnd'] = dataset.index.is_month_end
```


```{r, echo=FALSE, fig.align = "center"}
dataset_weekdays <- 
  py$dataset %>%
  group_by(Weekday)%>%
  summarise(across(c("First", "Second", "Third"), ~ mean(.x, na.rm = TRUE)))

dataset_weekdays$Weekday <- factor(dataset_weekdays$Weekday, labels = c('Monday','Tuesday','Wednesday','Thursday','Friday','Saturday','Sunday'))

dataset_weekdays_long <-
  melt(dataset_weekdays, id.vars = c("Weekday")
                       , measure.vars = c("First", "Second", "Third")
                       , variable.name = "Dose"
                       , value.name = "Count"
                    )

ggplot(data = dataset_weekdays_long, 
       aes(x=Weekday, 
           y=Count,
           fill=Count
           )
         ) +
  geom_col(position = "dodge") +
  geom_vline(xintercept = "Saturday", color = "red", linetype = "dotted") +
  facet_grid(Dose ~ .)
```

---

Are there other dependencies?

---

First of all, I am going to use Regression Machine Learning models: 

- [Decision Tree](https://scikit-learn.org/stable/modules/tree.html#regression) 

- [Random Forest](https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.RandomForestRegressor.html).

Which columns should I choose?

- "Weekday"

- "Year"

- "DayOfYear"

---
```{r python_functions, include=FALSE}
# function for preparing sets
source_python('prepare_sets.py')
# function for training models
source_python('train_model.py')
```

```{python my_columns, include=FALSE}
feature_columns = ["Weekday", "Year", "DayOfYear"]
```

```{python first, include=FALSE}
# The first column that I will predict is "First".
y_column = "First"
```

```{python train_all, echo=FALSE, render=pander::pander}
train_X, val_X, train_y, val_y = prepare_sets(dataset, feature_columns, y_column)

# DecisionTree
mae_dt, predictions_dt, model_dt = train_model(train_X, val_X, train_y, val_y, "DecisionTree", n_estimators=None)
# RandomForest
mae_rf, predictions_rf, model_rf = train_model(train_X, val_X, train_y, val_y, "RandomForest", n_estimators=500)

print("DecisionTree: ", 1 - mae_dt/dataset[y_column].mean())
print("RandomForest: ", 1 - mae_rf/dataset[y_column].mean())
```

```{python plot_all, echo=FALSE, fig.align = "center"}
plt.figure(figsize=(12,6))
sns.set_style("darkgrid")
plt.title('{} ({})'.format("Comparing", y_column))
# plot DecisionTree
sns.lineplot(data=predictions_dt, label='{} MAE:{}'.format("DecisionTree", round(mae_dt,0)))
# plot RandomForest
sns.lineplot(data=predictions_rf, label='{} MAE:{}'.format("RandomForest", round(mae_rf,0)))
# plot validation set
val_y.index=range(0,len(val_y))
sns.lineplot(data=val_y, label="Validation")
# add legend
plt.legend()
```

--- 

```{python second_train, include=FALSE}
y_column = "Second"
```

```{python, echo=FALSE}
<<train_all>>
<<plot_all>>
```

---

```{python third_train, include=FALSE}
y_column = "Third"
```

```{python, echo=FALSE}
<<train_all>>
<<plot_all>>
```

---

- The waves are recognisable.

- The extreme values are bigger than in real data.

---

# Question 3

Are there dependencies between academic year events and vaccination waves?

```{r, include=FALSE}
age_ord <- unique(bristol["age"])
bristol$Age <- which(age_ord == bristol$age)
```

```{r, include=FALSE}
y_column = "First"
```

---

```{r kmeans, echo=FALSE}
cls <- kmeans(x = data.matrix(bristol[,c("Age", y_column)]), centers = 5)
bristol$cluster <- as.character(cls$cluster)

ggplot() +
  geom_point(data = bristol, 
             mapping = aes_string(x = "age", 
                                  y = y_column, 
                                  colour = "cluster")) +
theme(axis.text.x = element_text(angle=90, hjust = 1))
```

---

```{r, echo=FALSE}
y_column = "Second"
<<kmeans>>
```

---

```{r, echo=FALSE}
y_column = "Third"
<<kmeans>>
```

```{r, include=FALSE}
dataset_student_long <- melt(filter(bristol, age=="18_24"), id.vars = c("date")
                       , measure.vars = c("First", "Second", "Third")
                       , variable.name = "Dose"
                       , value.name = "Count"
                    )
```

---

```{r, fig.height = 6.9, fig.width = 11.5, fig.align = "center", echo=FALSE}
#https://www.uwe.ac.uk/study/term-dates/2020-21-term-dates
#https://www.uwe.ac.uk/study/term-dates/2021-22-term-dates
#https://www.uwe.ac.uk/study/term-dates/2022-23-term-dates
ggplot(data = dataset_student_long,
       aes(x=date, y=Count, colour=Dose)) +
    geom_rect(aes(xmin = as.Date("2020-10-19"),
            xmax = as.Date("2021-01-22"),
                ymin = -Inf, ymax = Inf, fill = '20/21 Teaching Block 1: Autumn Term'), alpha = .2)+
    geom_rect(aes(xmin = as.Date("2021-02-08"),
            xmax = as.Date("2021-05-07"),
                ymin = -Inf, ymax = Inf, fill = '20/21 Teaching Block 2: Spring Term'), alpha = .2)+
    geom_rect(aes(xmin = as.Date("2021-05-24"),
            xmax = as.Date("2021-08-06"),
                ymin = -Inf, ymax = Inf, fill = '20/21 Teaching Block 3: Summer Term'), alpha = .2)+
    geom_rect(aes(xmin = as.Date("2021-09-27"),
            xmax = as.Date("2022-01-28"),
                ymin = -Inf, ymax = Inf, fill = '21/22 Teaching Block 1: Autumn Term'), alpha = .2)+
  geom_line() +
  facet_grid(Dose ~ .) +
  theme(axis.text.x = element_text(angle=90, hjust = 1)) +
  scale_x_date(date_breaks = "months" , date_labels = "%b-%y") +
  #scale_fill_brewer(palette = 'Dark2', name = 'Year') +
  scale_fill_hue(c=45, l=80)
```

# What is next?

- the UK data

- England data 

- deep learning and models considering time series.

- domain knowledge facts

# Thanks for your attention

To be continued... :) 