---
output:
  bookdown::pdf_document2: default
classoption: landscape
---
# Step 2 Ask something

```{r, include=FALSE}
#---------------------------------------
# pander -- beautiful string from chunk
#---------------------------------------
if(!require(pander)){install.packages("pander")}
library(pander)
```

I live in Bristol. What do I know about Bristol? 

* This city is a part of the UK, England, and South West. 
* There are two universities.
* The city rests every summer when students come back to their homes and works hardly elsewhen. 

So, I am interested in data about the UK, England, South West, and Bristol. 

**Question 0:** 
```{r q0, render=pander::pander, echo=FALSE}
"Are there dependencies between academic year events and vaccination waves? Does the vaccination depend on holidays?"
```

I was vaccinated by 

* the first dose on 8 August 2021,
* the second dose on 3 October 2021,
* the booster dose on 8 January 2022.

**Question 1:**
```{r q1, render=pander::pander, echo=FALSE}
"How many people got their jabs with me?"
```

I got the first and the second jabs on Sunday. There were fewer people in the vaccination centre. When I got the third jab on Saturday, there was a big queue. 

**Question 2:** 
```{r q2, render=pander::pander, echo=FALSE}
"When do people prefer to get a jab: weekdays or weekends/Saturdays or Sundays?"
```

**Question 3:** 
```{r q3, render=pander::pander, echo=FALSE}
"Is there something illogical in data?"
```