---
output:
  bookdown::pdf_document2: default
classoption: landscape
---
# Introduction {#Questions}

```{r, include=FALSE}
#---------------------------------------
# pander -- beautiful string from chunk
#---------------------------------------
if(!require(pander)){install.packages("pander")}
library(pander)
```

Hello. My name is Elena, and I am a Data Scientist. 

When I came to Bristol in May 2021, I decided to be vaccinated because, in my opinion, this is a safer way to live my normal life. So, I started my long research about Covid vaccination. And now I am ready to show you interesting facts.

## Questions

I live in Bristol. What do I know about Bristol? 

* This city is a part of the UK, England, and South West. 
* There are two universities.

$\text{\underline{Question 0:}}$
```{r q0, render=pander::pander, echo=FALSE}
"Are there dependencies between academic year events and vaccination waves?"
```

I was vaccinated by the first, the second, and booster doses on 8 August 2021, 3 October 2021, 8 January 2022, respectively. $\text{\underline{Question 1:}}$
```{r q1, render=pander::pander, echo=FALSE}
"How many people got their jabs with me?"
```

I got the first and the second jabs on Sunday. There were fewer people in the vaccination centre. When I got the third jab on Saturday, there was a big queue. 
$\text{\underline{Question 2:}}$
```{r q2, render=pander::pander, echo=FALSE}
"When do people prefer to get a jab: weekdays or weekends/Saturdays or Sundays?"
```

Finally, $\text{\underline{Question 3:}}$
```{r q3, render=pander::pander, echo=FALSE}
"Can I predict vaccination data?"
```

## Overall
I need to find datasets to answer my questions.

* Data is about the UK, England, South West, and Bristol.
* Data is about new people who got the jabs by dates.

And I was lucky to find the website ["Coronavirus (COVID-19) in the UK" ](https://coronavirus.data.gov.uk).
